Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review

### 2021-09-28 reports/manuscript.Rmd
Add tables row to index.html
Update manuscript and plots

---
title: "Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review"
author:
  - name: Taren Sanders
    affiliation: Institute for Positive Psychology and Education, Australian Catholic
      University
    affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
    orcid_id: 0000-0002-4504-6008
  - name: Philip Parker
    affiliation: Institute for Positive Psychology and Education, Australian Catholic
      University
    affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
    orcid_id: 0000-0002-4604-8566
  - name: Michael Noetel
    affiliation: School of Health and Behavioural Sciences, Australian Catholic University
    affiliation_url: https://www.acu.edu.au/about-acu/faculties-directorates-and-staff/faculty-of-health-sciences/school-of-behavioural-and-health-sciences
    orcid_id: 0000-0002-6563-8203
  - name: Chris Lonsdale
date: '`r format(Sys.Date())`'
repository_url: https://github.com/Motivation-and-Behaviour/screen_umbrella
bibliography: combined.bib
csl: nature.csl
params:
  render_img: true
  inc_ref: false
---

@@

@@

*Population*: To be eligible for inclusion, meta-analyses needed to include meta-analytic effect sizes for children or adolescents ( age 0-18 years). Meta-analyses containing data from adults and youth were included if meta-analytic effect sizes estimates specific to participants aged 18 years or less could be extracted. We excluded meta-analyses that only contained evidence gathered from adults (age >18 years).

@@

*Statistical Credibility*. We employed a statistical classification approach to grade the statistical credibility of the effect sizes in the literature. To be considered credible an effect needed to be derived from a combined sample of >1,000 and have non-significant tests of publication bias (i.e., Egger’s test and excess significance test). We performed these analyses, and tehrefore the review needed to provide usable study-level data in order to be included.

@@

```{r prisma, fig.cap="PRISMA Diagram"}
prisma$diag
```

@@

The most frequently reported exposures were `r knitr::combine_words(freq_exposures$out[1:4])`. The most frequently reported outcomes were `r knitr::combine_words(freq_outcomes$out[1:5])`. In most cases (`r paste0(sum(freq_combos$n==1), "/", nrow(effects_clean))`), there was only one exposure/outcome combination, with `r sum(freq_combos$n==2)` appearing twice, and `r sum(freq_combos$n>2)` three or more times. Full characteristics of the included studies are provided in Table \@ref(tab:desctable).

```{r desctable, layout="l-screen-inset shaded"}
reviews_tables[[1]]$table
```

@@

dualqualstr <- paste0(
  qual_str(n_high_dualquality, nrow(qual_data), " high ", suffix= "; "),
  qual_str(n_unclear_dualquality, nrow(qual_data), " unclear ", prefix = "n", suffix = ")"))

@@

The quality of the included meta-analyses was mixed. Most assessed heterogeneity `r qual_str(n_low_hetero, nrow(qual_data), " low ", suffix=" of studies)")`, reported the characteristics of the included studies `r qual_str(n_low_char, nrow(qual_data), " low ")`, and used a comprehensive and systematic search strategy `r qual_str(n_low_search, nrow(qual_data), " low ")`. Most reviews did not clearly report if their eligibility criteria was predefined `r qual_str(n_unclear_elig, nrow(qual_data), " unclear ")`. Many reviews also did not complete dual independent screening of abstracts and full text `r qual_str(n_high_dualscreen, nrow(qual_data), " high ")` or did not clearly report the method of screening `r qual_str(n_unclear_dualscreen, nrow(qual_data), " unclear ")`. Many reviews  also did not complete A similar trend was observed for dual independent quality assessment `r dualqualstr`. Overall, only `r all_low` meta-analyses were graded as *low* risk of bias on all criteria.

@@

There were `r n_edu_effect` unique effects associated with education outcomes, including general learning outcomes, literacy, numeracy, and science. Of these, `r n_edu_certain` met our criteria for statistical credibility and are described below (see Figure \@ref(fig:eduplot)). We removed `r n_edu_noindiv` for not providing individual study-level data, `r n_edu_samplesize` for having meta-analytic effects with samples < 1,000, and `r n_edu_faileggers` had a significant Egger's test or insufficient studies to conduct the test. No remaining studies showed evidence of excessive significance. Effects not meeting one or more of these standards are presented in the supplementary material.

```{r eduplot, fig.width=plots[[1]]$dims[[1]], fig.height=plots[[1]]$dims[[2]], fig.cap="Education outcomes"}
plots[[1]]$plot
```

Among the statistically credible effects, general screen use, TV viewing, and video games were all negatively associated with academic performance or general learning. E-books that included narration, as well as touch screen education interventions, and augmented reality interventions were positively associated with learning. General screen use was also negatively associated with literacy outcomes. However, if the screen involved co-viewing (e.g., watching with a parent) or the content was educational the association with literacy was positive. Numeracy outcomes were positively associated with screen-based mathematics interventions and video games with numeracy content.

@@

We identified `r n_health_effect` unique outcome-exposure combinations associated with health or health-related behaviour outcomes. `r xfun::numbers_to_words(n_health_certain, cap=TRUE)` of these meta-analytic associations met our criteria for credible evidence (see Figure \@ref(fig:healthplot)). We removed `r n_health_noindiv` for not providing individual study-level data, `r n_health_samplesize` for having meta-analytic effects with samples < 1,000, and `r n_health_faileggers` had a significant Egger's test or insufficient studies to conduct the test. No remaining studies showed evidence of excessive significance. Effects not meeting one or more of these standards are presented in the supplementary material.

```{r healthplot, fig.width=plots[[2]]$dims[[1]], fig.height=plots[[2]]$dims[[2]], fig.cap="Health and health-related behaviour outcomes"}
plots[[2]]$plot
```

@@

Television was negatively correlated with sleep duration, but only at the 95% confidence level (weak evidence). All forms of observed screen time (general, TV, and video games) were positively associated with body composition (e.g., BMI). Risky behaviour (e.g., unsafe sex) was also positively associated with social media and with the use of sexy media. General screen use was also associated with depression.

@@

Across the health outcomes, most (`r n_health_p999` of `r n_health_certain` effects) were statistically significant at the $P < 10^3$ level, with the remaining `r xfun::numbers_to_words(n_health_p95)` significant at the $P < .05$ level. However, all of the credible effects exhibited high levels of heterogeneity, with the lowest $I^2 = 75%$. Additionally, most effects were small, with the association between unhealthy food advertising and food intake the largest at $r = 0.25$. Most of the effect sizes (`r n_health_r2`/`r n_health_certain`) had an absolute value of $r < 0.2$.

@@

@@

```{r refs, results='asis'}
if (params$inc_ref) "# References"
```

>>>>

---
title: "Benefits and risks associated with children’s and adolescents’ interactions with electronic screens: An umbrella review"
author:
- name: Taren Sanders
  affiliation: Institute for Positive Psychology and Education, ACU
  affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
  orcid_id: 0000-0002-4504-6008
- name: Michael Noetel
  affiliation: School of Health and Behavioural Sciences, ACU
  affiliation_url: https://www.acu.edu.au/about-acu/faculties-directorates-and-staff/faculty-of-health-sciences/school-of-behavioural-and-health-sciences
  orcid_id: 0000-0002-6563-8203
- name: Philip Parker
  affiliation: Institute for Positive Psychology and Education, ACU
  affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
  orcid_id: 0000-0002-4604-8566
- name: Borja Del Pozo Cruz
- name: Stuart Biddle
- name: Rimante Ronto
- name: Ryan Hulteen
- name: Rhiannon Parker
- name: George Thomas
- name: Katrien De Cocker
- name: Jo Salmon
- name: Kylie Hesketh
- name: Nicole Weeks
- name: Hugh Arnott
- name: Emma Devine
- name: Roberta Vasconcellos
- name: Chris Lonsdale
  affiliation: Institute for Positive Psychology and Education, ACU
  affiliation_url: https://www.acu.edu.au/research/our-research-institutes/institute-for-positive-psychology-and-education
  orcid_id: 000-0002-2523-5565
date: '`r format(Sys.Date())`'
repository_url: https://github.com/Motivation-and-Behaviour/screen_umbrella
bibliography: combined.bib
csl: vancouver-brackets.csl
params:
  render_img: true
  inc_ref: false
---

@@

```{r, eval=params$inc_ref, results='markup'}

@@

*Population*: To be eligible for inclusion, meta-analyses needed to include meta-analytic effect sizes for children or adolescents (age 0-18 years). Meta-analyses containing data from adults and youth were included if meta-analytic effect sizes estimates specific to participants aged 18 years or less could be extracted. We excluded meta-analyses that only contained evidence gathered from adults (age >18 years).

@@

*Statistical Credibility*. We employed a statistical classification approach to grade the statistical credibility of the effect sizes in the literature. To be considered credible an effect needed to be derived from a combined sample of >1,000 and have non-significant tests of publication bias (i.e., Egger’s test and excess significance test). We performed these analyses, and therefore the review needed to provide usable study-level data in order to be included.

@@

```{r prisma, eval=params$render_img, fig.cap="PRISMA Diagram"}
prisma$diag
```

@@

The most frequently reported exposures were `r knitr::combine_words(freq_exposures$out[1:4])`. The most frequently reported outcomes were `r knitr::combine_words(freq_outcomes$out[1:5])`. In most cases (`r paste0(sum(freq_combos$n==1), "/", nrow(effects_clean))`), there was only one exposure/outcome combination, with `r sum(freq_combos$n==2)` appearing twice, and `r sum(freq_combos$n>2)` appearing three or more times. Full characteristics of the included studies are provided in Table \@ref(tab:desctable).

```{r desctable, eval=params$render_img, layout="l-screen-inset shaded"}
reviews_tables[[1]]$table
```

@@

dualqualstr <- paste0(
  qual_str(n_high_dualquality, nrow(qual_data), " high risk ", suffix= "; "),
  qual_str(n_unclear_dualquality, nrow(qual_data), " unclear  ", prefix = "n", suffix = ")"))

@@

The quality of the included meta-analyses was mixed. Most assessed heterogeneity `r qual_str(n_low_hetero, nrow(qual_data), " low risk ", suffix=" of studies)")`, reported the characteristics of the included studies `r qual_str(n_low_char, nrow(qual_data), " low risk ")`, and used a comprehensive and systematic search strategy `r qual_str(n_low_search, nrow(qual_data), " low risk ")`. Most reviews did not clearly report if their eligibility criteria was predefined `r qual_str(n_unclear_elig, nrow(qual_data), " unclear ")`. Many reviews also did not complete dual independent screening of abstracts and full text `r qual_str(n_high_dualscreen, nrow(qual_data), " high risk ")` or did not clearly report the method of screening `r qual_str(n_unclear_dualscreen, nrow(qual_data), " unclear ")`. A similar trend was observed for dual independent quality assessment `r dualqualstr`. Overall, only `r all_low` meta-analyses were graded as low risk of bias on all criteria.

@@

There were `r n_edu_effect` unique effects associated with education outcomes, including general learning outcomes, literacy, numeracy, and science. We removed `r n_edu_noindiv` effects that did not provide individual study-level data, `r n_edu_samplesize` effects with samples < 1,000, and `r n_edu_faileggers` with a significant Egger's test or insufficient studies to conduct the test. No remaining studies showed evidence of excessive significance. Effects not meeting one or more of these standards are presented in the supplementary material. The remaining `r n_edu_certain` met our criteria for statistical credibility and are described below (see Figure \@ref(fig:eduplot)).

```{r eduplot, eval=params$render_img, fig.width=plots[[1]]$dims[[1]], fig.height=plots[[1]]$dims[[2]], fig.cap="Education outcomes"}
plots[[1]]$plot
```

Among the statistically credible effects, general screen use, TV viewing, and video games were all negatively associated with academic performance or general learning. E-books that included narration, as well as touch screen education interventions, and augmented reality interventions were positively associated with learning. General screen use was negatively associated with literacy outcomes. However, if the screen involved co-viewing (e.g., watching with a parent) or the content was educational the association with literacy was positive. Numeracy outcomes were positively associated with screen-based mathematics interventions and video games with numeracy content.

@@

We identified `r n_health_effect` unique outcome-exposure combinations associated with health or health-related behaviour outcomes. We removed `r n_health_noindiv` effects that did not provide individual study-level data, `r n_health_samplesize` effects with samples < 1,000, and `r n_health_faileggers` effects with a significant Egger's test or insufficient studies to conduct the test. No remaining studies showed evidence of excessive significance. Effects not meeting one or more of these standards are presented in the supplementary material. The remaining `r n_health_certain` meta-analytic associations met our criteria for credible evidence and are described below (see also Figure \@ref(fig:healthplot)).

```{r healthplot, eval=params$render_img, fig.width=plots[[2]]$dims[[1]], fig.height=plots[[2]]$dims[[2]], fig.cap="Health and health-related behaviour outcomes"}
plots[[2]]$plot
```

@@

Television viewing was negatively correlated with sleep duration, but only at the 95% confidence level (weak evidence). All forms of screen time (general, TV, and video games) were positively associated with body composition (e.g., BMI). Risky behaviour (e.g., unsafe sex) was also positively associated with social media and with the use of sexy media. General screen use was also associated with depression.

@@

Across the health outcomes, most (`r n_health_p999` of `r n_health_certain` effects) were statistically significant at the $P < 10^3$ level, with the remaining `r xfun::numbers_to_words(n_health_p95)` significant at the $P < .05$ level. However, all of the credible effects exhibited high levels of heterogeneity, with the lowest $I^2 = 75\%$. Additionally, most effects were small, with the association between unhealthy food advertising and food intake the largest at $r = 0.25$. Most of the effect sizes (`r n_health_r2`/`r n_health_certain`) had an absolute value of $r < 0.2$.

@@

* [Descriptive Table](https://drive.google.com/file/d/1t5XvwY2kzsOsbZ2FxbGPsYEecZz5PgD1/view?usp=sharing)
* Supplemental tables (TBC)

@@

# References
